# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
library(plyr)
activity <- read.csv("activity.csv", header=TRUE)
activity[,2] <- as.Date(activity[,2], "%Y-%m-%d")

```

## What is mean total number of steps taken per day?
```{r fig.height=6, fig.width=7}
agg1 <- ddply(activity, .(date), summarize, total_steps=sum(steps), mean_steps=round(mean(steps),2), median_steps=median(steps))
hist(agg1$total_steps, breaks=10, xlab = "Total number of steps taken each day", main="Histogram of total number of steps taken each day")
agg1[,c("date","mean_steps","median_steps")]
```

## What is the average daily activity pattern?
```{r fig.height=6, fig.width=7}
agg2 <- ddply(activity, .(interval), summarize, avg_steps=round(mean(steps,na.rm=TRUE),2))
plot(agg2$interval, agg2$avg_steps, type="l", xlab="Interval",ylab="Average number of steps takenacross all days")
```
Below 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
agg2[agg2$avg_steps==max(agg2$avg_steps),]
```
## Imputing missing values
The total number of missing values in the dataset:
```{r}
length(which(is.na(activity)))
```

```{r fig.height=6, fig.width=7}
length(which(is.na(activity)))
activity2 <- activity
for (i in 1:nrow(activity2)) {
  if (is.na(activity2[i,"steps"])) {
    activity2[i,"steps"] <- agg2[agg2$interval == activity2[i,"interval"], "avg_steps"]
  }
}
agg3 <- ddply(activity2, .(date), summarize, total_steps=sum(steps), mean_steps=round(mean(steps),2), median_steps=median(steps))
hist(agg3$total_steps, breaks=10, xlab="Total number of steps taken each day", main="Histogram of total number of steps taken each day")
agg3[,c("date","mean_steps","median_steps")]
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity2$day_flag <- sapply(activity2$date, function(x) {if ((weekdays(x)=="Sunday") | (weekdays(x)=="Saturday")) {"weekend"} else {"weekday"}})
agg4 <- ddply(activity2, .(interval,day_flag), summarize, avg_steps=round(mean(steps,na.rm=TRUE),2))
agg4 <- transform(agg4, day_flag = factor(day_flag))
library(lattice)
xyplot(avg_steps ~ interval | day_flag, type="l", data=agg4,ylab="Number of steps", layout = c(1,2))
```
